%% Chapter 11 — Deployment Guide
\chapter{Deployment Guide}
\label{ch:deployment}

\section{Deployment Overview}

A complete MMS V2 deployment follows this order. Each step depends on the previous being complete.

\begin{table}[H]
\centering
\caption{Deployment Sequence}
\begin{tabular}{C{1cm}L{5cm}L{7.5cm}}
\toprule
\textbf{Step} & \textbf{Task} & \textbf{Reference} \\
\midrule
1  & Generate master key (ONCE) & Chapter~\ref{ch:deployment}, Section~\ref{sec:masterkey} \\
2  & Set up Firebase project & Deployment Package, Section 08 \\
3  & Set up Raspberry Pi & Deployment Package, Section 07 \\
4  & Deploy web application & Deployment Package, Section 09 \\
5  & Set up kiosk station & Deployment Package, Section 10 \\
6  & Assemble nodes & Chapter~\ref{ch:hardware}; Deployment Package, Section 05 \\
7  & Run hardware validation & Appendix~\ref{app:validation}; Deployment Package, Section 06 \\
8  & Provision nodes & Deployment Package, Section 11 \\
9  & Run integration tests & Deployment Package, Section 12 \\
10 & Issue cards to students & Deployment Package, Section 10 \\
11 & Pilot deployment (1 machine) & Deployment Package, Section 13 \\
12 & Full rollout (all machines) & Deployment Package, Section 13 \\
\bottomrule
\end{tabular}
\end{table}

\begin{notebox}
The full deployment package is in \texttt{v2/deployment/}. This chapter summarises the most critical steps. Refer to the deployment package for step-by-step instructions with exact commands.
\end{notebox}

\section{Master Key Generation}
\label{sec:masterkey}

The master key is a 32-byte random secret shared across all nodes and the kiosk. It must be generated exactly once, before any card is issued or any node is provisioned.

\begin{lstlisting}[style=bashstyle]
python3 -c "
import secrets
bs = secrets.token_bytes(32)
print('C array:')
print('  ' + ', '.join(f'0x{b:02X}' for b in bs))
print()
print('Python hex:')
print('  ' + bs.hex())
"
\end{lstlisting}

Store the output in:
\begin{itemize}[leftmargin=1.5em]
    \item A password manager (required)
    \item A printed copy in a sealed envelope in the lab safe (recommended)
    \item \textbf{Do not store in a shared Google Drive or any cloud storage in plaintext}
\end{itemize}

Use the C array form for \texttt{secrets.h} and \texttt{secrets\_station.h}. Use the hex form for \texttt{secrets.py}.

\begin{warnbox}
\textbf{Losing the master key means all existing cards must be replaced.} The key cannot be recovered from the cards or the firmware. If it is lost, generate a new key, re-flash all nodes, and re-issue all cards.
\end{warnbox}

\section{Node Provisioning Summary}

For each node:

\begin{enumerate}[leftmargin=1.5em]
    \item Copy \texttt{secrets.h.template} to \texttt{secrets.h}; fill in WiFi credentials and master key
    \item Set \texttt{MACHINE\_ID} and \texttt{MACHINE\_NAME} in \texttt{config.h}
    \item Flash firmware via Arduino IDE (Partition: Default 4MB with spiffs)
    \item Verify boot log on Serial Monitor (115200 baud)
    \item Confirm MQTT connect and config push from bridge
    \item Test card access: permitted card GRANTED, non-permitted card DENIED
    \item Complete provisioning checklist (Deployment Package, Section 11)
\end{enumerate}

\section{Critical Pre-Deployment Verification}

Before any node goes to a live machine, run these checks:

\begin{lstlisting}[style=bashstyle]
# 1. Bridge healthy
curl http://192.168.0.10:8080/health
# Expect: {"mqtt_connected": true, "fb_connected": true, ...}

# 2. Mosquitto reachable
mosquitto_pub -h 192.168.0.10 -t mms/test -m ping
# Expect: no error

# 3. Firestore machines seeded
firebase firestore:get /machines/1
# Expect: JSON with "Creality 3D Printer"

# 4. Web app accessible
# Open dashboard URL in browser; login; confirm 5 machines shown
\end{lstlisting}

\section{Secrets Files Reference}

\begin{table}[H]
\centering
\caption{Secret Files and Their Contents}
\begin{tabular}{L{6cm}L{8cm}}
\toprule
\textbf{File path} & \textbf{Contents} \\
\midrule
\texttt{v2/access\_node/secrets.h} & WiFi SSID/password, MQTT broker IP, master key (C array) \\
\texttt{v2/kiosk\_station/station\_writer/secrets\_station.h} & Master key (C array) \\
\texttt{v2/kiosk\_station/secrets.py} & Master key (Python hex bytes) \\
\texttt{v2/rpi\_bridge/.env} & MQTT credentials, Firebase RTDB URL \\
\texttt{v2/rpi\_bridge/service-account.json} & Firebase Admin SDK service account \\
\texttt{v2/web\_app/lab-access-nexus-main/.env} & Firebase client config (not a secret per se, but not committed) \\
\bottomrule
\end{tabular}
\end{table}

All of these are in \texttt{.gitignore}. Template files (e.g., \texttt{secrets.h.template}) are committed; actual secret files are not.
